---
title: "testing 1 hw"
author: "Hamed"
date: "2/9/2020"
output: html_document
---


#### a.	Read the data into a data frame in R

```{r ,warning=FALSE}
pdcalls=read.csv("jerseycitypdservicecalls2017csv.csv",header = T)
attach(pdcalls)
head(pdcalls,6)
```


#### b.	Find the number of variables and number of rows of data, The first number is for rows and the second for variables

```{r ,warning=FALSE}
dim(pdcalls)
```


#### c.	Inspect the data by printing out the first six rows.

```{r, warning=FALSE}
head(pdcalls,6)

```

# d.	How many rows are there with missing values?

```{r ,warning=FALSE}
sum(!complete.cases(pdcalls))

```

#   e.	Remove the rows with missing data

```{r ,warning=FALSE}
pdcalls <- pdcalls[complete.cases(pdcalls), ]
dim(pdcalls)

```

# f.	Check to see if there are any duplicate rows.  If there are duplicate rows then remove them.

```{r ,warning=FALSE}
which(duplicated(pdcalls)) #check for duplicates
pdcalls=pdcalls[!duplicated(pdcalls), ] #remove duplicates
dim(pdcalls)

```


#### g.	Sort the data by call type
```{r ,warning=FALSE}
head(pdcalls)
head(pdcalls[order(call.type),])

```


#### h.	Create a new data frame called calls911 by filtering the original dataset for the 911 calls.  Print out the first six rows and check if the filtering worked.  How many 911 calls were there?

```{r,warning=FALSE}
library(dplyr)
calls911=filter(pdcalls,call.type=='911')

head(calls911)
# Number of 911 calls
nrow(calls911)

```


#### i.	Create a new variable (column) called dispatchduration by subtracting time.received from time.dispatched.  Hint:  Use the mutate command in   dplyr. You also need to convert the format of time.received and time.dispatched using the strptime() command and then subtract. 

```{r,warning=FALSE}

library(dplyr)
pdcalls$time_R<-strptime(pdcalls$time.received,  # Apply strptime with timezone
                    format = "%m/%d/%Y %H:%M:%OS",
                    tz = "EST")


pdcalls$time_D<-strptime(pdcalls$time.dispatched,  # Apply strptime with timezone
                 format = "%m/%d/%Y %H:%M:%OS",
                 tz = "EST")


pdcalls$dispatchduration=difftime(pdcalls$time_D, pdcalls$time_R,
                                  units = "secs")

head(pdcalls)

```



#### j.	Now check if there are missing values in the newly created column and also check for dispatch durations that are negative or zero. This is garbage data so remove these rows.


```{r ,warning=FALSE}

#Missing values
sum(!complete.cases(pdcalls$dispatchduration))

#Negative values
nrow(pdcalls[pdcalls$dispatchduration<0,])

# Zero values
nrow(pdcalls[pdcalls$dispatchduration==0,])

# Remove negative and zero values
pdcalls_clean=pdcalls[pdcalls$dispatchduration>0,]
dim(pdcalls_clean)
head(pdcalls_clean)
```


#### k.	Find the average (mean) dispatch duration using the new variable
#(column) you created above.

```{r,warning=FALSE}
mean(pdcalls_clean$dispatchduration)
```


#### l.	Find the average (mean) dispatch duration by call type

```{r ,warning=FALSE}
library(doBy)
summaryBy(dispatchduration~call.type, data=pdcalls_clean, FUN=mean)

```



#### m.	How many rows contain the word GUNSHOTS in the call code description column.  Hint: Use the stringr package.  Install it if needed. Use the str_detect function in stringr.  Use the help to learn more about the function and how to use it.

```{r,warning=FALSE}
length(grep("GUNSHOTS", pdcalls$call.code.description))

```

#### n.	Now create a data frame called gunshotdata that has just the rows that contain the word GUNSHOTS in the call code description

```{r }
library(dplyr)
gunshotdata=dplyr::filter(pdcalls, grepl('GUNSHOTS', call.code.description))
head(gunshotdata)

```
