---
title: "Graphs and plots HW"
author: "HILLARY"
date: "1/31/2020"
output:
  word_document: default
  pdf_document: default
---

#### 1. Use the diamonds dataset that comes with R

```{r}
library(ggplot2)
data("diamonds")
attach(diamonds)
head(diamonds)
```

#### 2. Count the number of rows by clarity

```{r}
table(diamonds$clarity)
```

#### 3. What are the unique values for cut?

```{r}
unique(cut)
```


#### 4. Tabulate the frequency (no of diamonds) by cut

```{r}
table(diamonds$cut)

```


#### 5. Find the mean carat size by color

```{r}
library(dplyr)
diamonds %>%
  group_by(color) %>%
  summarise_at(vars(carat), funs(mean(., na.rm=TRUE)))

```

#### 6. Use a bar chart to find which cut has the most amount of diamonds

```{r}
require(lattice)
barchart(table(diamonds$cut),main="Number of diamonds per cut")

```

#### 7. Check the distribution of the carat size.  How can you describe the distribution?

- According to the plots the data is mostly found around the mean which is 0.79 but has a lot of outliers and thus the data is rightly skewed.  

```{r}
# Histogram plot
hist(diamonds$carat,main = "Histogram for carat size")

# Boxplot
boxplot(diamonds$carat,main="Boxplot for carat size")

#Density plot
d <- density(diamonds$carat)
plot(d, type="n", main="Density plot for carat size")
polygon(d, col="red", border="gray")

```

#### 8. Is there a relationship between carat size and price?  Check using a graph.
- From the graph the there is a positive slope showing carat size has a positive relationship with price.
```{r}
ggplot(diamonds, aes(carat,price)) + geom_point(alpha=0.3,col="#0000ff22", pch=16,cex=3)

```

#### 9. Which color has the maximum variability in the price?  Use a graph to find out.

-From the boxplots the color E shows higher variability in price.
```{r}
qplot(color,price,data=diamonds, geom = "boxplot")

```
