---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####The analysis include exploratoray data analysis and an ordinal logistic regression model.
```{r}

require("RCurl")
```
### Load the data
```{r}
#Raw dataset
newdata1<-getURL("https://raw.githubusercontent.com/watex95/R-FOR-DATA-SCIENCE/master/DATASETS/Gvault_survey_raw.csv")
df1<-read.csv(text=newdata1)
```

### EXPLORATORY DATA ANALYSIS
Load packages
```{r}
require("plotrix")
```

#### Lets explore the frequency of single choice questions
How frequently do you use GVault QDMS?
```{r}
table(df1$freq_use)
barplot(table(df1$freq_use),ylab = "responses",col = "light green",
      ,srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```





What is your role within GVault QDMS?
```{r}
table(df1$role)
barplot(table(df1$role),ylab = "responses",col = "light green",
        srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```






How would you describe the effectiveness of training you received?
```{r}
table(df1$training_effectiveness)
barplot(table(df1$training_effectiveness),ylab = "responses",col = "light green",srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```







Are you able to complete work in the system without 'help' or 'intervention'?
```{r}
require(plotrix)
mytable=table(df1$complete_without_help)
lbls <- paste(names(mytable), "\n", mytable, sep="")
pie3D(mytable, labels = lbls,explode=0.3)
```

Do you find it easy to search for and access documents in GVault QDMS?
```{r}
mytable=table(df1$easy_access_documents)
lbls <- paste(names(mytable), "\n", mytable, sep="")
pie3D(mytable, labels = lbls,explode=0.3)
```

How satisfied are you with the overall experience of GVault QDMS?
```{r}
mytable=table(df1$satisfaction)
lbls <- paste(names(mytable), "\n", mytable, sep="")
pie3D(mytable, labels = lbls,explode=0.3)
```

How would you describe your efficiency with GVault as compared to GDocs?
```{r}
table(df1$Gvault_efficiency)
barplot(table(df1$Gvault_efficiency),ylab = "responses",col = "light green",srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```






Do you find GVault has improved features and functionality compared to GDocs?
```{r}
table(df1$Gvault_improved)
barplot(table(df1$Gvault_improved),ylab="responses",
        col="light green",las=2, yaxt='n',cex.names = 0.65,cex.lab=0.65)


```







###Demographics analysis
What is your functional area?
```{r}
table(df1$functional_area)
barplot(table(df1$functional_area),ylab = "responses",col = "light green",srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```






What is your office location?
```{r}
table(df1$office_location)
barplot(table(df1$office_location),ylab = "responses",col = "light green",srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```









What is your Job Level?
```{r}
table(df1$job_level)
barplot(table(df1$job_level),ylab = "responses",col = "light green",
        srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```






How long have you worked at Gilead?
```{r}
table(df1$time_worked_Glead)
barplot(table(df1$time_worked_Glead),ylab = "responses",col = "light green",srt=45,las=2,yaxt='n',cex.names = 0.65,cex.lab=0.65)


```







###Ordinal Logistic Regression (OLR)
##### Load the Libraries
```{r}
require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)
```

For building this model, we will be using the polr command to estimate an ordered logistic regression. Then, we'll specify Hess=TRUE to let the model output show the observed information matrix from optimization which is used to get standard errors.

```{r}
m<-polr(df1$satisfaction ~ df1$freq_use + df1$role
          +df1$training_instructor_led + df1$training_web_based
          + df1$training_read+ df1$no_training+df1$training_effectiveness
          + df1$support_Gnet + df1$support_inapplication
          +df1$support_ref_doc + df1$support_SOP + df1$support_contacted
          +df1$support_IT + df1$complete_without_help
          +df1$easy_access_documents + df1$Gvault_efficiency
          +df1$Gvault_improved,data=df1,Hess=TRUE)
summary(m)
```


We see the usual regression output coefficient table including the value of each coefficient, standard errors, t values, estimates for the two intercepts, residual deviance and AIC. AIC is the information criteria. Lesser the better.

Now we'll calculate some essential metrics such as p-Value, CI, Odds ratio. The output displayed next are values: log odds(coefficients of estimation), t values are the t-distribution test statistic values, and finally the p-values are the probability of error.

```{r}
ctable <- coef(summary(m))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
```



#### Odds Ratios and their Confidence Intervals
```{r}
ci <- confint(m)
exp(cbind(OR = coef(m), ci))
```



####Interpretation at a glance: The effect of one predictor on the odds/likelihood of satisfaction
When Gvault is used atleast once a week versus atleast once a month, the odds of satisfaction being very satisfied are 0.9 greater than the odds of either "satisfied" , "slightly unsatisfied" or "very unsatisfied" combined. NOTE: The reference level selected was "atleast once a month" 

When the training is very effective versus effective training, the odds of a user being very satisfied are 4.38 greater than the odds of being either "satisfied" , "slightly unsatisfied" or "very unsatisfied" combined. NOTE: The reference level selected was "effective"

### Summary
Most of the predictor variables were not statistically significant in predicting the level of satisfaction with the overall experience of GVault QDMS. Training effectiveness (p-value=0.04688), Support for In application(p-value=0.00401), complete work without(p-value=0.000395) help,and whether Gvault improved features(p-value=0.0.00479) were the variables that significantly affected satisfaction of a user.

