---
title: "Homework 4.4 - Hitters dataset"
author: "Hamed"
date: "2/26/2020"
output: word_document
---

# We now use boosting to predict Salary in the Hitters data set in the ISLR package.

```{r, warning=FALSE}

library(ISLR)
Hitters=as.data.frame(Hitters)
dim(Hitters)
str(Hitters)

```


## (a) Remove the observations for whom the salary information is unknown, and then log-transform the salaries.

```{r, warning=FALSE}

# remove rows with missing values 
Hitters = Hitters[!is.na(Hitters$Salary),]

#log transform salary
Hitters$Salary = log(Hitters$Salary)


```


## (b) Create a training set consisting of 65% of the observations, and a test set consisting of the remaining observations.

```{r, warning=FALSE}
set.seed(123)
smp_size <-floor(0.65 * nrow(Hitters))
train_ind <-sample(seq_len(nrow(Hitters)), size = smp_size)
train.Hitters = Hitters[train_ind,]
dim(train.Hitters)
test.Hitters = Hitters[-train_ind,]
dim(test.Hitters)


```


## (c) Perform boosting using XGBoost on the training set with 1,000 trees for a range of values of the shrinkage parameter λ. Produce a plot with different shrinkage values on the x-axis and the corresponding training set MSE on the y-axis.

```{r, warning=FALSE}
library(xgboost)
library(caret) 
library(gbm)

# Ensure the R-studio can handle plots with large margins
par("mar")
par(mar=c(1,1,1,1))

#Grid search the shrinkage
set.seed(1)
shrinkGrid = seq(.1,.001,by=-.001)
shrinkGrid

#Define the MSE 
MSE = matrix(NA,nrow=length(shrinkGrid),ncol=2)

# Train xgboost model using the shrinkage
for(i in 1:length(shrinkGrid)){
  lambda = shrinkGrid[i]
  boost.hitters = gbm(Salary~.,data=train.Hitters,distribution="gaussian",n.trees=1000,shrinkage=lambda)
  boost.pred.train = predict(boost.hitters,train.Hitters,n.trees=1000)
  boost.pred.test = predict(boost.hitters,test.Hitters,n.trees=1000)
  train.MSE = mean((boost.pred.train - train.Hitters$Salary)^2)
  test.MSE = mean((boost.pred.test - test.Hitters$Salary)^2)
  MSE[i,1] = train.MSE
  MSE[i,2] = test.MSE
}
best.lambda = shrinkGrid[which.min(MSE[,2])]
best.lambda

# Plot the MSE against the shrinkage values for th train and test dataset
matplot(x=shrinkGrid,y=MSE,type="l",xlab="lambda",ylab="MSE")
legend("topright",legend = c("Train","Test"),col=c("black","red"),lty=c(1,2))
abline(v=best.lambda,col="blue",lty=2,lwd=.5)

#The best MSE 
print(paste0("Best Test MSE: ", MSE[which.min(MSE[,2]),2]))


```


## (d) Compare the test MSE of boosting to the test MSE that results from applying a linear regression approach. We shall apply two regression approaches; (ridge and lasso) 

```{r, warning=FALSE}
library(glmnet)

#split the dataset
x.train.Hitters = model.matrix(Salary~.,train.Hitters)[,-1]
x.test.Hitters = model.matrix(Salary~.,test.Hitters)[,-1]
y.train.Hitters = train.Hitters$Salary
y.test.Hitters = test.Hitters$Salary

#define lambda 
lambdaGrid = 10^seq(10,-2,length=100)

# Ridge Linear regression cross-validated
cv.ridge = cv.glmnet(x.train.Hitters,y.train.Hitters,alpha=0)
bestlam.ridge = cv.ridge$lambda.min

#Lasso linear regression cross-validated
cv.lasso = cv.glmnet(x.train.Hitters,y.train.Hitters,alpha=1)
bestlam.lasso = cv.lasso$lambda.min

#Ridge and lasso regression without cross-validation 
ridge.hitters = glmnet(x.train.Hitters,y.train.Hitters,alpha=0,lambda=lambdaGrid)
lasso.hitters = glmnet(x.train.Hitters,y.train.Hitters,alpha=1,lambda=lambdaGrid)

#Prediction
ridge.pred = predict(ridge.hitters,s=bestlam.ridge,newx=x.test.Hitters)
lasso.pred = predict(lasso.hitters,s=bestlam.lasso,newx=x.test.Hitters)

#Mean squared error 
ridge.MSE.test = mean((ridge.pred - y.test.Hitters)^2)
lasso.MSE.test = mean((lasso.pred - y.test.Hitters)^2)
print(paste0("Boost Test MSE: ", MSE[which.min(MSE[,2]),2]))
print(paste0("Ridge Test MSE: ", ridge.MSE.test))
print(paste0("Lasso Test MSE: ", lasso.MSE.test))

```


## (e) Which variables appear to be the most important predictors in the boosted model?

```{r, warning=FALSE}

summary(boost.hitters)

```


## (f) Now apply bagging to the training set. What is the test set MSE for this approach?

```{r, warning=FALSE}
  
library(randomForest)
bag.hitters = randomForest(Salary~.,data=train.Hitters,ntree=1000,mtry=19,importance=TRUE)

bag.pred = predict(bag.hitters,newdata=test.Hitters)
bag.MSE.test = mean((bag.pred - test.Hitters$Salary)^2)

print(paste0("Bagging Test MSE: ",bag.MSE.test))

# Bagging appears to have slightly outperformed boosting, ridge, and lasso.

```


