---
title: "Homework 4.1 - Auto dataset"
author: "Hamed"
date: "2/26/2020"
output: word_document
---
# In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the Auto data set in the ISLR library.

```{r,warning=FALSE}

#Load libraries and dataset
library(ISLR)
library(dplyr)
library(mlr)
library(caret)
library(ROCR)
library(kernlab)
library(e1071)
library(foreach)
library(doParallel)

Auto=as.data.frame(Auto)

```

## (a) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.

```{r,warning=FALSE}
med=median(Auto$mpg)
Auto$target=ifelse(Auto$mpg>med,'1','0')

# convert target to factor and drop the string variable for classification purposes

Auto$target=as.factor(as.character(Auto$target)) #convert to factor
class(Auto$target)
Auto=select(Auto,-starts_with('name')) #drop name

# Change to fatcor and perform Variable encoding on the origin variable because its a categorical variable
Auto$origin=as.factor(as.character(Auto$origin))
Auto=createDummyFeatures(Auto, cols = "origin")
str(Auto)

```

## (b) Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validated training dataset error associated with different values of this parameter. Comment on your results.

```{r,warning=FALSE}
# split the dataset
set.seed(123)
smp_size<- floor(0.85 * nrow(Auto))
train_ind<-sample(seq_len(nrow(Auto)), size = smp_size)
train.Auto<-Auto[train_ind, ]
test.Auto<-Auto[-train_ind, ]

# Fit the model
SVM1=tune(svm,target~.,data = train.Auto,ranges=list(cost=c(.01,.02,.05,
                .1,.2,.5,1,2,5,10)),kernel='linear',
            trControl = trainControl("cv", number = 5))

SVM1$best.model
summary(SVM1) #The error reduces as the model cross validates towards 10 folds


```

## (c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma (sigma) and degree and cost. Comment on your results.

```{r,warning=FALSE}
# RADIAL KERNEL
SVM2 = tune(svm ,target~.,data=train.Auto,ranges=list(
  cost=c(.01,.02,.05,.1,.2,.5,1,2,5,10),gamma=c(.001,.002,.005,
        .01,.02,.05,.1,.2,.5,1,2,5,10)),kernel='radial',
        trControl = trainControl("cv", number = 10),
        preProcess = c("center","scale"))

SVM2$best.model
summary(SVM2)


```


```{r,warning=FALSE}
# POLYNOMIAL

SVM3 <- train(target ~., data = train.Auto, method = "svmPoly",
               trControl = trainControl("cv", number = 10),
               tuneGrid = expand.grid(C=c(.01,.02,.05,.1,.2,.5,1,2,5,10)
              ,degree=c(1:5),scale=c(0.01:1)),
               preProcess = c("center","scale"),
               tuneLength = 4)

SVM3$finalModel #displays cost , error, degree and scale of the model
SVM3$results #displays the accuracy of the model crossvalidated 


```

