---
title: "Homework 4.3 - Carseats dataset"
author: "Hamed"
date: "2/26/2020"
output: word_document
---

# Use the Carseats data set in ISLR package

```{r,warning=FALSE}
library(ISLR)
library(tree)

carseats=as.data.frame(Carseats)
str(carseats)
dim(carseats)

```

## (a) Split the data set into a training set and a test set.

```{r,warning=FALSE}
set.seed(123)
smp_size <-floor(0.80 * nrow(carseats))
smp_size

train_ind <-sample(seq_len(nrow(carseats)), size = smp_size)
train.carseats = carseats[train_ind,]
dim(train.carseats)
test.carseats = carseats[-train_ind,]
dim(test.carseats)

```


## (b) Fit a regression tree to the training set by selecting an appropriate target variable to predict. Plot the tree, and interpret the results. What test MSE do you obtain?

```{r,warning=FALSE}
#Regression model tree 
tree.carseats = tree(Sales~.,data=train.carseats)
summary.carseats = summary(tree.carseats)
summary.carseats

#Plot the tree
plot(tree.carseats)
text(tree.carseats,pretty=0)

#Prediction
tree.pred = predict(tree.carseats, newdata=test.carseats)
head(tree.pred)

#Mean squared error
mean((tree.pred-test.carseats$Sales)^2)

```

## (c) Use cross-validation in order to determine the optimal level of tree complexity. 

```{r,warning=FALSE}
#Cross validation
cv.carseats = cv.tree(tree.carseats)
names(cv.carseats)
cv.carseats

#plot the deviance against the size of the tree model
plot(cv.carseats$size, cv.carseats$dev,type="b")
points(which.min(cv.carseats$dev),cv.carseats$dev[which.min(cv.carseats$dev)],col="red",pch=19,cex=1.25)

#Prune and Plot the pruned cross validated tree
prune.carseats = prune.tree(tree.carseats,best = which.min(cv.carseats$dev))
plot(prune.carseats)
text(prune.carseats,pretty=0)

#Prediction
prune.pred = predict(prune.carseats, test.carseats)
head(prune.pred)

#Mean squared error
mean((prune.pred-test.carseats$Sales)^2)

```

## (d) Use the bagging approach to do the prediction. What test MSE do you  obtain? Which variables are most important? Plot the variable importance.

```{r,warning=FALSE}

library(randomForest)
set.seed(1)

#Bagging
bag.carseats = randomForest(Sales~.,data=train.carseats,mtry=10,importance=T)
bag.carseats

#Prediction and mean squared error (MSE)
yhat.bag = predict(bag.carseats,newdata=test.carseats)
mean((yhat.bag - test.carseats$Sales)^2)

#Variable importance
importance(bag.carseats)
varImpPlot(bag.carseats)

```

# (e) Use random forest to do the prediction. What test MSE do you obtain? 

```{r,warning=FALSE}
#Value of m=5
rf5.carseats = randomForest(Sales~.,data=train.carseats,mtry=5,importance=T)
rf5.carseats

#Prediction and MSE
yhat5.rf = predict(rf5.carseats,newdata=test.carseats)
mean((yhat5.rf-test.carseats$Sales)^2)


```

## Try various values of m, the number of variables considered at each split and describe the effect of m on the error rate.

```{r,warning=FALSE}

#Value of m=3
rf3.carseats = randomForest(Sales~.,data=train.carseats,mtry=3,importance=T)
rf3.carseats

#Prediction and MSE
yhat3.rf = predict(rf3.carseats,newdata=test.carseats)
mean((yhat3.rf-test.carseats$Sales)^2)

```


## Which variables are most important. Plot the variable importance. 

```{r,warning=FALSE}

par(mfrow=c(2,2))
varImpPlot(rf5.carseats)
varImpPlot(rf3.carseats)

```
